---
title: "threshold"
output: html_document
date: "2022-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r library}
library(tidyverse)
library(ggplot2)
```

## model

```{r }
#####CONSTANTS####

# number of individuals n

n<-30

#number of tasks m

m<-2

#Probability exponent
eta<-10 # in paper range from 1 to 30
#all tasks same demand rate
demandRate<-0.6 
# if all demand differtn stimulusRate<-vector(mode="numeric", length=m)

#reduction in stimulus if all individuals 
#in colony do the task
alpha<-50

#quit probability
tau<-0.2

corrStep       <- 200 #number of time steps for calculation of correlation 


threshMean<-10 #forr both tasks
threshSD<-0.1
#time
timesteps<-10000
ThreshM        <- c(10, 10) #population threshold means 
ThreshSD       <- ThreshM * 0 #population threshold standard deviations 
InitialStim    <- c(0, 0) #intital vector of stimuli
StimRates      <- c(0.6, 0.6) #vector of stimuli increase rates  
threshSlope    <- 30 #exponent parameter for threshold curve shape  
alpha          <- m #efficiency of task performance
quitP          <- 0.2 #probability of quitting task once active



```

```{r}

#### FUNCTIONS####

#stimulus update function
stimFunc<-function(stimVec,activityMat,demandRate,alpha,n){
  output<-stimVec+demandRate-alpha*(colSums(activityMat)/n)
  output[output<0]<-0
  return(output)
}

probabilityfunc<-function(threshMat,stimVec,n,m,eta){
  probMat<-matrix(data = 0, nrow=n,ncol=m)
  for (i in 1:m) {
    s<-stimVec[i]
    denominator<-s^eta+threshMat[,i]^eta
    P<-(s^eta)/denominator
    probMat[,i]<-P
  }
  return(probMat)
}
```


```{r}
####SYSTEM SETUP####

#activity matrix

activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)


# internal  thresholds (same for all tasks)

thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)

```

```{r}
#### DATA RECORDS####
stimTrack<-matrix(data=0,nrow=timesteps,ncol=m)
activityTrack<-matrix(data=0,nrow=timesteps,ncol=n)
workersTrack<-matrix(data=0,nrow=timesteps,ncol=m)

```

```{r}
####TIMESTEPPING####

for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data
  stimTrack[t,]<-stimulusVec
  activityTrack[t,]<-rowSums(t(t(activityMat)*1:m))
  workersTrack<-colSums(activityMat)
  }
  

```

```{r}
#### plot data####
timeElapsed<-1:timesteps
stimTrackGraph<-cbind(timeElapsed,stimTrack) %>% as.data.frame()


ggplot(data=stimTrackGraph)+
  geom_line(aes(x=timeElapsed,y=V2))+
  geom_line(aes(x=timeElapsed,y=V3))
```

```{r}
#workers 

#task performance frequency profile
workers<-1:n
task1<-colSums(activityTrack==1)/timesteps
task2<-colSums(activityTrack==2)/timesteps

freqProfile<-cbind(workers,task1,task2) %>% as.data.frame()
task1<-colSums(activityTrack==1)/colSums(activityTrack>0)
task2<-colSums(activityTrack==2)/colSums(activityTrack>0)

freqProfile2<-cbind(workers,task1,task2) %>% as.data.frame()

ColonyAvgs<-freqProfile %>% summarise(mean(task1),sd(task1),mean(task2),sd(task2))

ColonyAvgs

freqProfile2 %>% summarise(mean(task1),sd(task1),mean(task2),sd(task2))


ggplot(data=freqProfile2 %>% pivot_longer(cols = c(2,3), names_to = "task",values_to = "frequency"))+
  geom_histogram(aes(x=frequency,y=..density..,fill=as.factor(task)), position="identity", alpha=0.4)
```


### try to determine fate of one individual

```{r}
roleDeterminismFunc<-function(n,m,threshMean,threshSD,individualThresh,reps,timesteps){
  timeProportion<-matrix(data=0,nrow=reps,ncol=m)
  for (r in 1:reps) {
    thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)
thresholdMat[1,]<-c(individualThresh,2*threshMean-individualThresh)

activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)

antTrack<-vector(mode="numeric", length=timesteps)
for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data
  antTrack[t]<-sum(activityMat[1,]*1:m)
  }
task1<-sum(antTrack==1)
task2<-sum(antTrack==2)

timeProportion[r,]<-c(sum(task1)/timesteps,sum(task2)/timesteps)
  }
  return(timeProportion)
}

roleDeterminismFunc(n=10,m=2,threshMean=10, threshSD=0.1, individualThresh = 10,reps=100,timesteps=10000) %>% colMeans()

```

##DETERMINISM###

What do we want to compare to?
baseline: all ants are the same
baseline: all ants are random
--what are the effects of fixing the ant?
--what are the effects of the inherent stochasticity?
--what do we want to understand about the system

test: run all ants the same on long timescales--even out the stochasticity?

What parts of variation is caused by the thresholds vs other sources of stochasticity?

paper figure 7 but from the perspective of the individual ant
--which sources of stochasticity are most responsible for the spread?--fix thresholds 

create nested problems: start from baselines
isolate  each source of stochasticity

task neglect--only at larger group sizes task neglect is less of a problem

Merlijn:ask questions about individual variation  in the limit not the short term--easier starting point: what is the influence of other ants?

analogy: living in an apartment--how  much does the choice  of co-residents influence which tasks you specialize on?

pull timestep choice  from paper

problem to address: runaway task demand? should we ask the question only under conditions where greoups can  manage?

follow up paper-


game plan--consider baseline cases and special cases so we don't overinterpret the findings

make sure groups don't fail, figure out simulations
answer questions with groups with uniform thresholds

consider completely random assemblages


What determines behavior?
individual traits
group traits
random chance

Meaningful question!



```{r}
individualThresh<-c(threshMean-threshMean*threshSD,threshMean+threshMean*threshSD)
```

```{r}
n<-3
m<-2
reps<-200
timesteps<-5000 ## run this for longer
```

```{r}
timeProportion<-matrix(data=0,nrow=reps,ncol=m)
thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)
thresholdMat[1,]<-individualThresh  
for (r in 1:reps) {

activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)

antTrack<-vector(mode="numeric", length=timesteps)
for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data
  antTrack[t]<-sum(activityMat[1,]*1:m)
  }
task1<-sum(antTrack==1)
task2<-sum(antTrack==2)

timeProportion[r,]<-c(sum(task1)/timesteps,sum(task2)/timesteps)
  }

```
```{r}
antFate<-timeProportion/rowSums(timeProportion)
hist(antFate[,2])
hist(antFate[,1])
```


```{r}
n<-10
reps<-200
timesteps<-10000


#constants from figure 6 of paper
m<-2
threshMean<-10
threshSD<-0.1
eta<-7
tau<-0.2
demandRate<-0.6

```

```{r}
timeProportion<-matrix(data=0,nrow=reps,ncol=m)
antThresh_1<-matrix(data=0,nrow = reps,ncol=n)
antThresh_2<-matrix(data=0,nrow = reps,ncol=n)
antTimeTask1<-matrix(data=0,nrow=reps,ncol=n)
antTimeTask2<-matrix(data=0,nrow=reps,ncol=n)

antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
antTimeTask2_track<-matrix(data=0,nrow=1,ncol=n)
  for (r in 1:reps) {
    thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)


activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)

antTrack<-vector(mode="numeric", length=timesteps)
antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
antTimeTask2_track<-matrix(data=0,nrow=1,ncol=n)
for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data
  antTrack[t]<-sum(activityMat[1,]*1:m)
  antTimeTask1_track<-antTimeTask1_track+activityMat[,1]
  antTimeTask2_track<-antTimeTask2_track+activityMat[,2]
  }
task1<-sum(antTrack==1)
task2<-sum(antTrack==2)

timeProportion[r,]<-c(sum(task1)/timesteps,sum(task2)/timesteps)

antThresh_1[r,]<-thresholdMat[,1]
antThresh_2[r,]<-thresholdMat[,2]
antTimeTask1[r,]<-antTimeTask1_track
antTimeTask2[r,]<-antTimeTask2_track
  }

```

```{r}
antFate<-timeProportion/rowSums(timeProportion)
hist(antFate[,1])
hist(antFate[,2])
```
##principal component analysis?? 
consider difference between thresholds or ratios  of thresholds to each other or colony averages
increase colony size---compare 30 ant colonies--what  do we measure? expect less of a spread? or less dependence on threshold 2?
Vs THRESHOLD
```{r}
library(ggplot2)
library(dplyr)
T1_vec<-antThresh_1 %>% as.vector()
T2_vec<-antThresh_2 %>% as.vector()
Task1_vec<-antTimeTask1 %>% as.vector()
Task2_vec<-antTimeTask2 %>% as.vector()
Task1_vec<-Task1_vec/timesteps
Task2_vec<-Task2_vec/timesteps
Task1_colonyavg_deviation<-T1_vec-rowMeans(antThresh_1)
Task2_colonyavg_deviation<-T2_vec-rowMeans(antThresh_2)
df<-cbind(T1_vec,T2_vec,Task1_vec,Task2_vec,Task1_colonyavg_deviation,Task2_colonyavg_deviation) %>% as.data.frame()
colnames(df)<-c("task1_threshold","task2_threshold","time_on_task1","time_on_task2","task1_threshold_colonyavg_deviation","task2_threshold_colonyavg_deviation")



ggplot(df)+
  geom_point(aes(x=task1_threshold,y=time_on_task1),col="blue",alpha=0.5)+
  geom_point(aes(x=task2_threshold,y=time_on_task2),col="red",alpha=0.5)

ggplot(df)+
  geom_point(aes(x=task1_threshold,y=time_on_task1),col="blue",alpha=0.5)+
  geom_smooth(aes(x=task1_threshold,y=time_on_task1),method="lm")

ggplot(df)+
  geom_point(aes(x=task2_threshold,y=time_on_task1),col="blue",alpha=0.5)+
  geom_smooth(aes(x=task2_threshold,y=time_on_task1),method="lm")

ggplot(df)+
  geom_point(aes(x=task1_threshold_colonyavg_deviation,y=time_on_task1),col="blue",alpha=0.5)+
  geom_smooth(aes(x=task1_threshold_colonyavg_deviation,y=time_on_task1),method="lm")

ggplot(df)+
  geom_point(aes(x=task2_threshold_colonyavg_deviation,y=time_on_task1),col="blue",alpha=0.5)+
  geom_smooth(aes(x=task2_threshold_colonyavg_deviation,y=time_on_task1),method="lm")


lm_task1_vs_task1_thresh<-lm(time_on_task1~task1_threshold,data=df)
summary(lm_task1_vs_task1_thresh)
lm_task1_vs_task1_thresh_dev<-lm(time_on_task1~task1_threshold_colonyavg_deviation,data=df)
summary(lm_task1_vs_task1_thresh_dev)
lm_task1_vs_task2_thresh<-lm(time_on_task1~task2_threshold,data=df)
summary(lm_task1_vs_task2_thresh)


rgl::plot3d(x=df$task1_threshold_colonyavg_deviation, y=df$task2_threshold_colonyavg_deviation, z=df$time_on_task1)
```

What to compare each time? 

take the end?

```{r}
#attempt to replicate figure 6a
n_vec<-c(1,2,4,6,8,12,16,32,100)
reps<-100
timesteps<-2000

antTimeTask1<-matrix(data=0,nrow=reps,ncol=n)
T1_performancefreq_all<-matrix(data = NA,nrow=1,ncol=4) %>% as.data.frame()
colnames(T1_performancefreq_all) <- c("iteration","ant","task1_freq","num_ants")
antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
for (w in 1:length(n_vec)) {
  n<-n_vec[w]
  antTimeTask1<-matrix(data=0,nrow=reps,ncol=n)
  for (r in 1:reps) {
    thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)


activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)

antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
antTimeTask2_track<-matrix(data=0,nrow=1,ncol=n)
for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data

  antTimeTask1_track<-antTimeTask1_track+activityMat[,1]

  }


antTimeTask1[r,]<-antTimeTask1_track



  }
  antTimeTask1_df<-as.data.frame(antTimeTask1)
  antTimeTask1_df$iteration<-c(1:reps)
  addendum<- antTimeTask1_df %>% pivot_longer(cols=1:n, names_to = "ant",values_to = "task1_freq") %>% mutate(num_ants=n)
  T1_performancefreq_all<-rbind(T1_performancefreq_all,addendum)
  
}

T1_performancefreq_all<-T1_performancefreq_all %>% na.omit()
T1_performancefreq_all<-T1_performancefreq_all %>% mutate(task1_prop=task1_freq/timesteps)

ggplot(T1_performancefreq_all)+
  geom_point(mapping=aes(x=num_ants,y=task1_prop))+
  geom_boxplot(aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),alpha=0.2)
ggplot(T1_performancefreq_all %>% filter(num_ants<25))+
  geom_jitter(mapping=aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),width=0.5,height=0)+
  geom_boxplot(aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),alpha=0.2)

plot_df<-T1_performancefreq_all %>% mutate(x_coord<-num_ants)
```

```{r}
#attempt to replicate figure 6a
n_vec<-c(1,2,3,4,6,8,10,12,16,32,100)
reps<-100
timesteps<-10000

antTimeTask1<-matrix(data=0,nrow=reps,ncol=length(n_vec))
T1_performancefreq_all<-matrix(data = NA,nrow=1,ncol=4) %>% as.data.frame()
colnames(T1_performancefreq_all) <- c("iteration","ant","task1_freq","num_ants")
antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
for (w in 1:length(n_vec)) {
  n<-n_vec[w]
  antTimeTask1<-matrix(data=0,nrow=reps,ncol=n)
  for (r in 1:reps) {
    thresholdMat<-matrix(data=rnorm(n*m,mean = threshMean,sd=threshSD*threshMean),nrow=n,ncol=m)


activityMat<-matrix(data=0,nrow=n,ncol=m)

#active(1) or inactive (0)

activity<-vector(mode="numeric", length=n)

#stimulus
stimulusVec<-vector(mode="numeric", length=m)

antTimeTask1_track<-matrix(data=0,nrow=1,ncol=n)
antTimeTask2_track<-matrix(data=0,nrow=1,ncol=n)
for (t in 1:timesteps) {
  #update probabilities
  probabilities<-probabilityfunc(threshMat=thresholdMat, stimVec = stimulusVec,n=n,m=m, eta=eta)
  
  #update individuals
  activity<-rowSums(activityMat)
  quits<-rbinom(n=n,size=1,prob=tau)
  for (i in 1:n) {
    #active individuals may become inactive
    if (activity[i]==1&&quits[i]==1) {
        activityMat[i,]<-matrix(data=0,nrow=1,ncol=m)
    }
    #inactive individuals may become active 
    if (activity[i]==0) {
      seq<-sample(1:m,m)
      for (j in 1:m) {
        activity[i]<-sum(activityMat[i,])
        if (activity[i]==0) {
          taskNum<-seq[j]
          taskProb<-probabilities[i,taskNum]
          activityMat[i,taskNum]<-rbinom(n=1,size=1,prob=taskProb)
        }
      }
    }
    }
  
  #change stimulus from activity in the past timestep
  stimulusVec<-stimFunc(stimulusVec,activityMat,demandRate,alpha,n) 
  
  #record data

  antTimeTask1_track<-antTimeTask1_track+activityMat[,1]

  }


antTimeTask1[r,]<-antTimeTask1_track



  }
  antTimeTask1_df<-as.data.frame(antTimeTask1)
  antTimeTask1_df$iteration<-c(1:reps)
  addendum<- antTimeTask1_df %>% pivot_longer(cols=1:n, names_to = "ant",values_to = "task1_freq") %>% mutate(num_ants=n)
  T1_performancefreq_all<-rbind(T1_performancefreq_all,addendum)
  
}

T1_performancefreq_all<-T1_performancefreq_all %>% na.omit()
T1_performancefreq_all<-T1_performancefreq_all %>% mutate(task1_prop=task1_freq/timesteps)

ggplot(T1_performancefreq_all)+
  geom_point(mapping=aes(x=num_ants,y=task1_prop))+
  geom_boxplot(aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),alpha=0.2)
ggplot(T1_performancefreq_all %>% filter(num_ants<25))+
  geom_jitter(mapping=aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),width=0.5,height=0)+
  geom_boxplot(aes(x=num_ants,y=task1_prop,color=as.factor(num_ants)),alpha=0.2)

plot_df<-T1_performancefreq_all %>% mutate(x_coord<-num_ants)
```

